---
title: "Amy Mimic Thesis"
author: "Amy.Wu"
date: "2024-11-27"
output: github_document
---

# -----------------------------------------------------------
# Setup: Install and Load Required Libraries
# -----------------------------------------------------------
```{r}
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Vector of packages needed for this project
required_pkgs <- c("dplyr", "readr", "ggplot2", "survival", 
                   "randomForest", "stringr")

# Install any packages that aren't already installed, then load them
for (pkg in required_pkgs) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg, dependencies = TRUE)
    library(pkg, character.only = TRUE)
  }
}
```
# -----------------------------------------------------------
# Aim One: Investigate the relationship between ICU stay duration 
#          and wound healing outcomes in diabetic patients, 
#          identifying key predictors of impaired healing.
# -----------------------------------------------------------

# -----------------------------------------------------------
# Section 1: Load Data from Local CSV Files
# -----------------------------------------------------------
```{r}
# Define file paths (update these paths as necessary)
data_dir <- "/Users/wuxiaoyu/Desktop/9419 Master's Essay/mimic-iv-3.1"
patients_file           <- file.path(data_dir, "hosp", "patients.csv")
icustays_file           <- file.path(data_dir, "icu", "icustays.csv")
diagnoses_icd_file      <- file.path(data_dir, "hosp", "diagnoses_icd.csv")
labevents_file          <- file.path(data_dir, "hosp", "labevents.csv")
procedureevents_file    <- file.path(data_dir, "icu", "procedureevents.csv")
microbiologyevents_file <- file.path(data_dir, "hosp", "microbiologyevents.csv")
d_labitems_file         <- file.path(data_dir, "hosp", "d_labitems.csv")
d_icd_procedures_file   <- file.path(data_dir, "hosp", "d_icd_procedures.csv")

# Read CSV files using readr::read_csv for faster performance
patients            <- read_csv(patients_file)
icustays            <- read_csv(icustays_file)
diagnoses_icd       <- read_csv(diagnoses_icd_file)
labevents           <- read_csv(labevents_file)
procedureevents     <- read_csv(procedureevents_file)
microbiologyevents  <- read_csv(microbiologyevents_file)
d_labitems          <- read_csv(d_labitems_file)
d_icd_procedures    <- read_csv(d_icd_procedures_file)
```
# -----------------------------------------------------------
# Section 2: Define the Diabetic Cohort
# -----------------------------------------------------------
```{r}
library(dplyr)
library(stringr)

# -----------------------------------------------------------
# Step 1: Filter Adult Patients (Aged ≥18 Years)
# -----------------------------------------------------------
adult_patients <- patients %>%
  filter(anchor_age >= 18) %>%
  select(subject_id, anchor_age) %>%
  distinct()

# -----------------------------------------------------------
# Step 2: Identify Diabetic Patients Using ICD Codes
# -----------------------------------------------------------
diabetic_patients <- diagnoses_icd %>%
  filter(str_starts(icd_code, "250") |  # ICD-9: 250.xx
         str_starts(icd_code, "E08") |  # ICD-10: E08
         str_starts(icd_code, "E09") |
         str_starts(icd_code, "E10") |
         str_starts(icd_code, "E11") |
         str_starts(icd_code, "E13")) %>%
  # Remove subject_id to avoid duplicating it in the join with icustays
  select(hadm_id) %>%
  distinct()

# -----------------------------------------------------------
# Step 3: Extract ICU Stays for Diabetic Patients
# -----------------------------------------------------------
diabetic_icu_stays <- icustays %>%
  inner_join(diabetic_patients, by = "hadm_id") %>%  # Merge using hadm_id
  filter(los >= 0.5) %>%  # Adjust filter to ≥12 hours (0.5 days)
  select(subject_id, hadm_id, stay_id, los) %>%
  distinct()

# -----------------------------------------------------------
# Step 4: Identify Wound-Related Diagnoses and Procedures
# -----------------------------------------------------------
wound_icd_procedures <- d_icd_procedures %>%
  filter(str_detect(long_title, regex("wound|ulcer|debridement|dressing|skin|abscess", ignore_case = TRUE))) %>%
  select(icd_code) %>%
  distinct()

wound_procedures <- procedureevents %>%
  filter(itemid %in% wound_icd_procedures$icd_code) %>%
  select(hadm_id, itemid, starttime) %>%  
  distinct()

# -----------------------------------------------------------
# Step 5: Identify Infection-Related Laboratory Markers
# -----------------------------------------------------------
infection_markers <- d_labitems %>%
  filter(str_detect(label, regex("infection|WBC|CRP|sepsis|culture", ignore_case = TRUE))) %>%
  select(itemid) %>%
  distinct()

infection_labs <- labevents %>%
  filter(itemid %in% infection_markers$itemid) %>%
  select(hadm_id, itemid, value, valuenum, flag) %>%
  distinct()

# -----------------------------------------------------------
# Step 6: Merge Data and Apply Exclusion Criteria
# -----------------------------------------------------------
filtered_patients <- diabetic_icu_stays %>%
  left_join(wound_procedures, by = "hadm_id") %>%  # LEFT JOIN to include all ICU stays
  left_join(infection_labs, by = "hadm_id") %>%       # LEFT JOIN for infection labs
  filter(!is.na(los)) %>%  # Ensure ICU LOS is recorded
  distinct()

# -----------------------------------------------------------
# Step 7: Count Final Filtered Cohort
# -----------------------------------------------------------
num_filtered_patients <- filtered_patients %>%
  distinct(hadm_id) %>%
  count()

# Print the final cohort count
print(num_filtered_patients)

# Preview the final filtered dataset
head(filtered_patients)
```
# -----------------------------------------------------------
# Section 3: Define Wound Healing Outcomes
# -----------------------------------------------------------
```{r}
# (A) Identify Infection-Related Lab Tests
# Get unique lab test descriptions from d_labitems
lab_test_list <- d_labitems %>%
  select(itemid, label) %>%
  distinct()

# Save and inspect the lab test list if needed
write_csv(lab_test_list, "lab_test_list.csv")

# Filter for infection-related terms (e.g., "infection", "WBC", "CRP", "sepsis", "culture")
infection_markers <- lab_test_list %>%
  filter(str_detect(label, regex("infection|WBC|CRP|sepsis|culture", ignore_case = TRUE)))
write_csv(infection_markers, "infection_markers.csv")

# Vector of infection marker item IDs
infection_lab_items <- infection_markers$itemid

# (B) Filter labevents for infection markers and join with ICU stay info
labevents_with_stay_id <- labevents %>%
  inner_join(icustays, by = c("subject_id", "hadm_id")) %>%
  select(subject_id, stay_id, hadm_id, itemid, value, valuenum, flag)

# Filter for infection markers
wound_healing_labs <- labevents_with_stay_id %>%
  filter(itemid %in% infection_lab_items) %>%
  select(subject_id, stay_id, itemid, value, valuenum, flag)
head(wound_healing_labs)

# (C) Identify Wound-Related Procedures
# Extract unique procedure descriptions from d_icd_procedures
procedure_list <- d_icd_procedures %>%
  select(icd_code, long_title) %>%  # Adjust column names if needed
  distinct()
write_csv(procedure_list, "procedure_list.csv")

# Filter for wound-related keywords in procedure descriptions
wound_icd_procedures <- procedure_list %>%
  filter(str_detect(long_title, regex("wound|ulcer|debridement|dressing|skin|abscess", ignore_case = TRUE)))
write_csv(wound_icd_procedures, "wound_icd_procedures.csv")

# Get wound-related ICD codes and filter procedureevents
wound_icd_codes <- wound_icd_procedures$icd_code
wound_procedures <- procedureevents %>%
  filter(itemid %in% wound_icd_codes) %>%  # Assuming itemid contains the procedure codes
  select(subject_id, stay_id, hadm_id, itemid, starttime, value)
write_csv(wound_procedures, "wound_procedures.csv")
head(wound_procedures)

# (D) Extract Wound-Related Microbiology Events
# List and inspect unique specimen types
unique_specimen_types <- microbiologyevents %>%
  select(spec_type_desc) %>%
  distinct()
write_csv(unique_specimen_types, "unique_specimen_types.csv")

# Filter for wound-related specimen types
wound_specimens <- microbiologyevents %>%
  filter(str_detect(spec_type_desc, regex("wound|tissue|abscess|drainage", ignore_case = TRUE)))
write_csv(wound_specimens, "wound_specimens.csv")
head(wound_specimens)

# Further filter for wound-related test names (e.g., "culture", "bacteria", etc.)
wound_tests <- wound_specimens %>%
  filter(str_detect(test_name, regex("culture|bacteria|susceptibility|stain", ignore_case = TRUE))) %>%
  select(subject_id, hadm_id, spec_type_desc, test_name, org_name, ab_name, comments)
write_csv(wound_tests, "wound_tests.csv")
head(wound_tests)

# Extract and save wound test results (comments and results)
wound_results <- wound_tests %>%
  select(subject_id, hadm_id, test_name, org_name, ab_name, comments)
write_csv(wound_results, "wound_results.csv")
head(wound_results)
```

# -----------------------------------------------------------
# Section 4: Combine Data into a Unified Dataset
# -----------------------------------------------------------
```{r}
# Ensure wound_healing_labs includes the full join keys
wound_healing_labs <- labevents %>%
  inner_join(icustays, by = c("subject_id", "hadm_id")) %>%
  select(subject_id, hadm_id, stay_id, itemid, value, valuenum, flag) %>%
  filter(itemid %in% infection_lab_items)  # infection_lab_items defined earlier

# Ensure wound_procedures includes the full join keys
wound_procedures <- procedureevents %>%
  filter(itemid %in% wound_icd_codes) %>%  # wound_icd_codes defined from wound_icd_procedures
  select(subject_id, hadm_id, stay_id, itemid, starttime, value)

icu_wound_data <- diabetic_icu_stays %>%  # diabetic_icu_stays already contains subject_id, hadm_id, stay_id, los
  # (A) Join Demographic Information (Covariates) from patients
  left_join(
    select(patients, subject_id, anchor_age, gender, anchor_year, anchor_year_group, dod),
    by = "subject_id"
  ) %>%
  
  # (B) Join ICU Unit Information (excludes 'los' to preserve the original column)
  left_join(
    select(icustays, subject_id, hadm_id, stay_id, first_careunit, last_careunit),
    by = c("subject_id", "hadm_id", "stay_id")
  ) %>%
  
  # (C) Join Infection-Related Lab Results (Exposure/Outcome: Infection Marker Lab Values)
  left_join(
    wound_healing_labs,
    by = c("subject_id", "hadm_id", "stay_id")
  ) %>%
  
  # (D) Join Wound-Related Procedures (Exposure: Clinical Interventions)
  left_join(
    wound_procedures,
    by = c("subject_id", "hadm_id", "stay_id"),
    suffix = c("", ".proc")
  ) %>%
  
  # (E) Join Wound Healing Outcomes (Outcome: Wound Test Results)
  left_join(
    wound_results,  # wound_results includes subject_id, hadm_id, test_name, comments, etc.
    by = c("subject_id", "hadm_id")
  ) %>%
  
  # (F) Join Comorbid Conditions (Covariate)
  left_join(
    diagnoses_icd %>%
      filter(!(str_starts(icd_code, "250") |
               str_starts(icd_code, "E08") |
               str_starts(icd_code, "E09") |
               str_starts(icd_code, "E10") |
               str_starts(icd_code, "E11") |
               str_starts(icd_code, "E13"))) %>%
      group_by(subject_id, hadm_id) %>%
      summarise(comorbidities = paste(unique(icd_code), collapse = "; "), .groups = "drop"),
    by = c("subject_id", "hadm_id")
  ) %>%
  
  # (G) Create Additional Outcome/Exposure Variables
  mutate(prolonged_stay = if_else(los > 7, 1, 0))

# Replace any empty column names with a placeholder ("unknown")
icu_wound_data <- icu_wound_data %>%
  rename_with(~ if_else(. == "", "unknown", .))

# Optionally, save the combined dataset to a CSV file
write_csv(icu_wound_data, "icu_wound_data_combined.csv")

# Preview the final combined dataset
head(icu_wound_data)

```

# -----------------------------------------------------------
# Section 5: Data Preprocessing and Cleaning
# -----------------------------------------------------------
```{r}
library(dplyr)
# (A) Remove Duplicate Rows
icu_wound_data <- icu_wound_data %>%
  distinct()

# (B) Consolidate and Clean 'value' Columns
# If both 'value.x' and 'value.y' exist, convert them to numeric and coalesce;
# otherwise, convert an existing 'value' column or create a placeholder.
if (all(c("value.x", "value.y") %in% names(icu_wound_data))) {
  icu_wound_data <- icu_wound_data %>%
    mutate(across(c(value.x, value.y),
                  ~ as.numeric(if_else(. %in% c("N/A", ""),
                                       NA_character_,
                                       as.character(.))))) %>%
    mutate(value = coalesce(value.x, value.y))
} else if ("value" %in% names(icu_wound_data)) {
  icu_wound_data <- icu_wound_data %>%
    mutate(value = as.numeric(if_else(value %in% c("N/A", ""),
                                      NA_character_,
                                      as.character(value))))
} else {
  warning("No 'value' columns found in icu_wound_data. Creating a placeholder 'value' column as NA.")
  icu_wound_data <- icu_wound_data %>%
    mutate(value = NA_real_)
}

# (C) Remove Rows with Missing Critical Values ('value' or 'los')
icu_wound_data <- icu_wound_data %>%
  filter(!is.na(value) & !is.na(los))

# Optionally, Impute Any Remaining Missing 'value' Entries with the Mean
icu_wound_data <- icu_wound_data %>%
  mutate(value = if_else(is.na(value), mean(value, na.rm = TRUE), value))

# Verify the Data Structure, Summary, and a Preview of the Data
str(icu_wound_data)
summary(icu_wound_data)
head(icu_wound_data)
```

# -----------------------------------------------------------
# Section 6: Descriptive Statistics and Visualization
# -----------------------------------------------------------
```{r}
# Print summary statistics for ICU length of stay and infection marker values
summary(icu_wound_data$los)
summary(icu_wound_data$value)

# Create a grouped summary by prolonged ICU stay status (0 = No, 1 = Yes)
prolonged_summary <- icu_wound_data %>%
  group_by(prolonged_stay) %>%
  summarise(
    count = n(),                         # Number of observations
    mean_los = mean(los, na.rm = TRUE),    # Mean length of stay
    mean_value = mean(value, na.rm = TRUE),# Mean infection marker value
    .groups = "drop"                      # Ungroup after summarising
  )
print(prolonged_summary)

# -------------------------------
# Visualization
# -------------------------------

# Histogram of ICU Length of Stay
ggplot(icu_wound_data, aes(x = los)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(
    title = "Distribution of ICU Stay Duration",
    x = "Length of Stay (days)",
    y = "Frequency"
  )

# Boxplot of Infection Marker Values by Prolonged ICU Stay Status
ggplot(icu_wound_data, aes(x = factor(prolonged_stay), y = value)) +
  geom_boxplot(fill = "orange") +
  labs(
    title = "Infection Marker Values by Prolonged ICU Stay",
    x = "Prolonged Stay (0 = No, 1 = Yes)",
    y = "Infection Marker Value"
  )
```

# -----------------------------------------------------------
# Section 7: Regression Analysis
# -----------------------------------------------------------
```{r}
# Linear regression: Predict infection marker value using ICU LOS and prolonged stay flag
reg_model <- lm(value ~ los + prolonged_stay, data = icu_wound_data)
summary(reg_model)

# Plot regression results: Scatter plot with fitted regression line
ggplot(icu_wound_data, aes(x = los, y = value)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Regression: ICU Stay Duration vs Infection Marker Value",
       x = "ICU Length of Stay (days)",
       y = "Infection Marker Value")
```

# -----------------------------------------------------------
# Section 8: Kaplan-Meier Survival Analysis
# -----------------------------------------------------------
```{r}
# Here we simulate wound healing times.
# Replace this simulation with actual healing data if available.
set.seed(123)  # For reproducibility
icu_wound_data <- icu_wound_data %>%
  mutate(
    healing_time = los + sample(1:5, n(), replace = TRUE),
    healed = if_else(value < 5, 1, 0)  # Define healing based on infection marker threshold
  )

# Fit Kaplan-Meier survival model
km_fit <- survfit(Surv(healing_time, healed) ~ prolonged_stay, data = icu_wound_data)

# Plot Kaplan-Meier survival curves
plot(km_fit, col = c("blue", "red"), lwd = 2,
     xlab = "Time to Wound Healing (days)",
     ylab = "Survival Probability",
     main = "Kaplan-Meier Curve: Time to Wound Healing")
legend("topright", legend = c("Short ICU Stay", "Prolonged ICU Stay"),
       col = c("blue", "red"), lwd = 2)
```

# -----------------------------------------------------------
# Section 9: Random Forest Feature Importance
# -----------------------------------------------------------
```{r}
# Prepare data for random forest analysis by selecting relevant variables and removing NAs
rf_data <- icu_wound_data %>%
  select(value, los, prolonged_stay) %>%
  na.omit()

# Train the random forest model (predicting infection marker value)
rf_model <- randomForest(
  value ~ los + prolonged_stay,
  data = rf_data,
  importance = TRUE,
  ntree = 500
)

# Print and plot feature importance
print(importance(rf_model))
varImpPlot(rf_model)
```
# -----------------------------------------------------------
# Section 10: Identifying Key Predictors of Impaired Healing
# -----------------------------------------------------------
```{r}
# Examine the range of infection marker values
icu_wound_data %>% 
  summarise(
    min_value = min(value, na.rm = TRUE),
    max_value = max(value, na.rm = TRUE)
  )
# Expected output: min_value = 0, max_value = 961

# Create a binary outcome for impaired healing using a threshold of 100.
# Here, values >= 100 are considered indicative of impaired healing.
icu_wound_data <- icu_wound_data %>%
  mutate(impaired_healing = if_else(value >= 100, 1, 0))

# Check the distribution of the new outcome variable
print(table(icu_wound_data$impaired_healing))

# Fit a logistic regression model to assess predictors of impaired healing
logit_model <- glm(impaired_healing ~ los + prolonged_stay, 
                   data = icu_wound_data, 
                   family = binomial)

# Display the summary of the logistic regression model
summary(logit_model)

# Calculate odds ratios and 95% confidence intervals for the predictors
exp_coef <- exp(coef(logit_model))
conf_int <- exp(confint(logit_model))
odds_ratios <- data.frame(
  Estimate = exp_coef, 
  Lower_CI = conf_int[, 1], 
  Upper_CI = conf_int[, 2]
)
print(odds_ratios)
```

# -----------------------------------------------------------
# Section 11: Cox Proportional Hazards Model for Time to Healing
# -----------------------------------------------------------
```{r}
# Ensure the survival package is loaded
library(survival)

# Fit the Cox proportional hazards model using healing time and the event indicator (healed)
cox_model <- coxph(Surv(healing_time, healed) ~ los + prolonged_stay, 
                   data = icu_wound_data)

# Summarize the Cox model to review hazard ratios (HR) and p-values
summary(cox_model)
```

```{r}
# Count the total number of unique patients in the dataset
total_unique_patients <- patients %>%
  distinct(subject_id) %>%
  count()

# Print the result
print(total_unique_patients)
```

```{r}
library(dplyr)
library(stringr)

#-----------------------------------------------------------
# 1. Total Unique Patients Admitted to Both Hospital and ICU
#-----------------------------------------------------------
# Assuming 'patients' holds all hospital admissions and 'icustays' holds ICU admissions,
# we get the intersection of subject IDs in both datasets.
total_hosp_ids <- unique(patients$subject_id)
total_icu_ids  <- unique(icustays$subject_id)
total_unique_patients <- length(intersect(total_hosp_ids, total_icu_ids))
print(paste("Total unique patients (hospital & ICU):", total_unique_patients))

#-----------------------------------------------------------
# 2. Adult Patients (Aged ≥18 Years)
#-----------------------------------------------------------
adult_patients <- patients %>%
  filter(anchor_age >= 18) %>%
  select(subject_id, anchor_age) %>%
  distinct()
adult_patient_count <- nrow(adult_patients)
print(paste("Adult patients (>= 18 years):", adult_patient_count))

#-----------------------------------------------------------
# 3. Diabetic Patients Based on ICD Codes (All Admissions)
#-----------------------------------------------------------
# Here we extract diabetic admissions from the diagnoses_icd table.
# Note: This assumes a column "subject_id" is available in diagnoses_icd.
diabetic_patients_all <- diagnoses_icd %>%
  filter(str_starts(icd_code, "250") |   # ICD-9: 250.xx
         str_starts(icd_code, "E08") |   # ICD-10: E08
         str_starts(icd_code, "E09") |
         str_starts(icd_code, "E10") |
         str_starts(icd_code, "E11") |
         str_starts(icd_code, "E13")) %>%
  select(subject_id) %>%   # If subject_id exists here
  distinct()
diabetic_patient_count <- nrow(diabetic_patients_all)
print(paste("Diabetic patients (documented diagnosis):", diabetic_patient_count))

#-----------------------------------------------------------
# 4. Diabetic Patients with ICU Stays (LOS ≥ 12 hours)
#-----------------------------------------------------------
# This step merges diabetic patients with ICU stays (and filters for LOS ≥ 0.5 days)
diabetic_patients_icustays <- diagnoses_icd %>%
  filter(str_starts(icd_code, "250") |  
         str_starts(icd_code, "E08") |
         str_starts(icd_code, "E09") |
         str_starts(icd_code, "E10") |
         str_starts(icd_code, "E11") |
         str_starts(icd_code, "E13")) %>%
  select(hadm_id) %>% 
  distinct()

diabetic_icu_stays <- icustays %>%
  inner_join(diabetic_patients_icustays, by = "hadm_id") %>% 
  filter(los >= 0.5) %>%      # ≥12 hours (0.5 days)
  select(subject_id, hadm_id, stay_id, los) %>%
  distinct()
diabetic_icu_stays_count <- nrow(diabetic_icu_stays)
print(paste("Diabetic ICU stays (LOS >= 12 hours):", diabetic_icu_stays_count))

#-----------------------------------------------------------
# 5. Exclusions Due to Incomplete Data or Inconsistent Admission Timestamps
#-----------------------------------------------------------
# For example, if you have a column indicating missing key demographic data,
# you can compute the number excluded. Here we assume that diabetic_icu_stays had some
# missing or inconsistent records that were later removed.
# (In the code below, filtered_patients is the post-exclusion dataset.)

filtered_patients <- diabetic_icu_stays %>%
  left_join(
    procedureevents %>%
      filter(itemid %in% (d_icd_procedures %>%
                            filter(str_detect(long_title, regex("wound|ulcer|debridement|dressing|skin|abscess",
                                                                   ignore_case = TRUE))) %>%
                            select(icd_code) %>%
                            distinct() %>%
                            pull(icd_code))),
    by = "hadm_id") %>%
  left_join(
    labevents %>%
      filter(itemid %in% (d_labitems %>%
                            filter(str_detect(label, regex("infection|WBC|CRP|sepsis|culture", ignore_case = TRUE))) %>%
                            select(itemid) %>%
                            distinct() %>%
                            pull(itemid))),
    by = "hadm_id") %>%
  filter(!is.na(los)) %>%
  distinct()

excluded_incomplete <- diabetic_icu_stays_count - nrow(filtered_patients)
print(paste("Patients excluded due to incomplete data/inconsistent timestamps:", excluded_incomplete))

#-----------------------------------------------------------
# 6. Exclusions Due to Lack of Wound-Related Complications
#-----------------------------------------------------------
# If you further require evidence of wound-related complications (i.e., either a wound-related
# procedure or infection-related lab marker must be present), then compute:
wound_evidence <- filtered_patients %>%
  filter(!is.na(starttime) | !is.na(valuenum))
  # Adjust based on which columns indicate wound evidence
wound_evidence_count <- nrow(wound_evidence)
excluded_wound <- nrow(filtered_patients) - wound_evidence_count
print(paste("Patients excluded due to no evidence of wound-related complications:", excluded_wound))

#-----------------------------------------------------------
# 7. Final Study Sample
#-----------------------------------------------------------
final_sample_count <- wound_evidence_count  # or use filtered_patients if you don't further restrict
print(paste("Final study sample size:", final_sample_count))
```

```{r}
# Print the head of each dataset

cat("Patients data:\n")
print(head(patients))

cat("\nICU Stays data:\n")
print(head(icustays))

cat("\nDiagnoses ICD data:\n")
print(head(diagnoses_icd))

cat("\nLab Events data:\n")
print(head(labevents))

cat("\nProcedure Events data:\n")
print(head(procedureevents))

cat("\nMicrobiology Events data:\n")
print(head(microbiologyevents))

cat("\nD Lab Items data:\n")
print(head(d_labitems))

cat("\nD ICD Procedures data:\n")
print(head(d_icd_procedures))

library(dplyr)
library(tibble)

# Create a named list of your datasets
datasets <- list(
  patients = patients,
  icustays = icustays,
  diagnoses_icd = diagnoses_icd,
  labevents = labevents,
  procedureevents = procedureevents,
  microbiologyevents = microbiologyevents,
  d_labitems = d_labitems,
  d_icd_procedures = d_icd_procedures
)

# For each dataset, get a comma-separated string of column names
head_names_table <- tibble(
  File = names(datasets),
  Column_Headers = sapply(datasets, function(df) paste(names(df), collapse = ", "))
)

# Print the summary table
print(head_names_table)
```