---
title: "Mimic Thesis"
author: "Amy.Wu"
date: "2024-11-27"
output: github_document
---

```{r}
## Set CRAN mirror
options(repos = c(CRAN = "https://cloud.r-project.org"))

## Install required libraries if not already installed
install.packages("dplyr")
install.packages("readr")
install.packages("ggplot2")
install.packages("survival")
install.packages("randomForest")
install.packages("readr")
install.packages("stringr")

## Load libraries
library(dplyr)
library(readr)
library(ggplot2)
library(survival)
library(randomForest)
library(readr)
library(stringr)
```

# Aim One: investigate the relationship between ICU stay duration and wound healing outcomes in diabetic patients, identifying key predictors of impaired healing;

# Section 1: Load Data from Local CSV Files
```{r}
# Load relevant tables into R
patients <- read.csv("/Users/wuxiaoyu/Desktop/9419 Master's Essay/mimic-iv-3.1/hosp/patients.csv")
icustays <- read.csv("/Users/wuxiaoyu/Desktop/9419 Master's Essay/mimic-iv-3.1/icu/icustays.csv")
diagnoses_icd <- read.csv("/Users/wuxiaoyu/Desktop/9419 Master's Essay/mimic-iv-3.1/hosp/diagnoses_icd.csv")
labevents <- read.csv("/Users/wuxiaoyu/Desktop/9419 Master's Essay/mimic-iv-3.1/hosp/labevents.csv")
procedureevents <- read.csv("/Users/wuxiaoyu/Desktop/9419 Master's Essay/mimic-iv-3.1/icu/procedureevents.csv")
microbiologyevents <- read.csv("/Users/wuxiaoyu/Desktop/9419 Master's Essay/mimic-iv-3.1/hosp/microbiologyevents.csv")
```
# Section 2: Define the Cohort
```{r}
# Identify Diabetic Patients

## Filter for diabetes-related ICD codes (ICD-9: 250.xx, ICD-10: E08-E13)
diabetic_patients <- diagnoses_icd %>%
  filter(
    str_starts(icd_code, "250") | # ICD-9
    str_starts(icd_code, "E08") | # ICD-10
    str_starts(icd_code, "E09") |
    str_starts(icd_code, "E10") |
    str_starts(icd_code, "E11") |
    str_starts(icd_code, "E13")
  ) %>%
  select(subject_id) %>%
  distinct()

## View the first few diabetic patients
head(diabetic_patients)

# Extract ICU Stays

## Filter ICU stays for diabetic patients
diabetic_icu_stays <- icustays %>%
  inner_join(diabetic_patients, by = "subject_id") %>%
  select(subject_id, stay_id, los) # Keep only relevant columns

## View the first few ICU stays
head(diabetic_icu_stays)

```
# Section 3: Define Wound Healing Outcomes
```{r}
#Look for Wound Healing Indicators Using labevents, procedureevents, or microbiologyevents:
## Load the d_labitems table

# labevents
d_labitems <- read.csv("/Users/wuxiaoyu/Desktop/9419 Master's Essay/mimic-iv-3.1/hosp/d_labitems.csv")

## Extract unique lab test descriptions
lab_test_list <- d_labitems %>%
  select(itemid, label) %>%
  distinct()

## View the first few rows
head(lab_test_list)

## Save the list to a CSV file for easier inspection
write.csv(lab_test_list, "lab_test_list.csv", row.names = FALSE)

## Filter lab test descriptions for infection-related terms
infection_markers <- lab_test_list %>%
  filter(str_detect(label, regex("infection|WBC|CRP|sepsis|culture", ignore_case = TRUE)))

## View the filtered list
head(infection_markers)

## Save the filtered list to a CSV file for detailed review
write.csv(infection_markers, "infection_markers.csv", row.names = FALSE)

## Vector of identified infection marker item IDs
infection_lab_items <- infection_markers$itemid

## Filter labevents for infection markers
wound_healing_labs <- labevents %>%
  filter(itemid %in% infection_lab_items) %>%
  select(subject_id, hadm_id, itemid, value, valuenum, flag) # Use hadm_id instead of stay_id

## View the first few rows
head(wound_healing_labs)

## Join labevents with icustays to add stay_id
labevents_with_stay_id <- labevents %>%
  inner_join(icustays, by = c("subject_id", "hadm_id")) %>% # Match on subject_id and hadm_id
  select(subject_id, stay_id, hadm_id, itemid, value, valuenum, flag)

## View the first few rows
head(labevents_with_stay_id)

## Filter for infection markers
wound_healing_labs <- labevents_with_stay_id %>%
  filter(itemid %in% infection_lab_items) %>%
  select(subject_id, stay_id, itemid, value, valuenum, flag)

## View the first few rows
head(wound_healing_labs)

# procedureevents

## Load d_icd_procedures table
d_icd_procedures <- read.csv("/Users/wuxiaoyu/Desktop/9419 Master's Essay/mimic-iv-3.1/hosp/d_icd_procedures.csv")

## View the structure of the table
head(d_icd_procedures)

## Extract unique ICD procedure descriptions
procedure_list <- d_icd_procedures %>%
  select(icd_code, long_title) %>% # Replace `long_title` with the appropriate column name if necessary
  distinct()

## View the first few rows
head(procedure_list)

## Save to a CSV file for manual inspection
write.csv(procedure_list, "procedure_list.csv", row.names = FALSE)

## Filter for wound-related procedures
wound_icd_procedures <- procedure_list %>%
  filter(str_detect(long_title, regex("wound|ulcer|debridement|dressing|skin|abscess", ignore_case = TRUE)))

## View the filtered list
head(wound_icd_procedures)

## Save the filtered list to a CSV file for review
write.csv(wound_icd_procedures, "wound_icd_procedures.csv", row.names = FALSE)

## Extract ICD codes for wound-related procedures
wound_icd_codes <- wound_icd_procedures$icd_code

## Filter procedureevents for wound-related procedures
wound_procedures <- procedureevents %>%
  filter(itemid %in% wound_icd_codes) %>% # Use `itemid` as the column for ICD codes
  select(subject_id, stay_id, hadm_id, itemid, starttime, value) # Select relevant columns

## View the first few rows
head(wound_procedures)

## Save the filtered procedures to a CSV file
write.csv(wound_procedures, "wound_procedures.csv", row.names = FALSE)

# microbiologyevents

## List unique specimen types
unique_specimen_types <- microbiologyevents %>%
  select(spec_type_desc) %>%
  distinct()

## View the unique specimen types
head(unique_specimen_types)

## Save to a CSV for manual inspection
write.csv(unique_specimen_types, "unique_specimen_types.csv", row.names = FALSE)

## Filter for wound-related specimen types
wound_specimens <- microbiologyevents %>%
  filter(str_detect(spec_type_desc, regex("wound|tissue|abscess|drainage", ignore_case = TRUE)))

## View the first few rows
head(wound_specimens)

## Save to a CSV for detailed review
write.csv(wound_specimens, "wound_specimens.csv", row.names = FALSE)

## List unique tests for wound specimens
unique_wound_tests <- wound_specimens %>%
  select(test_name) %>%
  distinct()

## View the unique wound tests
head(unique_wound_tests)

## Save to a CSV for manual inspection
write.csv(unique_wound_tests, "unique_wound_tests.csv", row.names = FALSE)

## Example: Filter for specific wound-related tests
wound_tests <- wound_specimens %>%
  filter(str_detect(test_name, regex("culture|bacteria|susceptibility|stain", ignore_case = TRUE))) %>%
  select(subject_id, hadm_id, spec_type_desc, test_name, org_name, ab_name, comments)

## View the first few rows
head(wound_tests)

## Save to a CSV for detailed review
write.csv(wound_tests, "wound_tests.csv", row.names = FALSE)

##  Extract comments and results for review
wound_results <- wound_tests %>%
  select(subject_id, hadm_id, test_name, org_name, ab_name, comments)

##  View the first few rows
head(wound_results)

##  Save to a CSV for detailed review
write.csv(wound_results, "wound_results.csv", row.names = FALSE)
```
# Section 4: Combine Data
```{r}
# Combine ICU stay data with wound-related lab, procedure, and microbiology results.

## Merge ICU stays with lab events and procedures
icu_wound_data <- diabetic_icu_stays %>%
  left_join(wound_healing_labs, by = c("subject_id", "stay_id")) %>%
  left_join(wound_procedures, by = c("subject_id", "stay_id")) %>%
  left_join(wound_results, by = c("subject_id", "hadm_id")) %>%
  mutate(prolonged_stay = ifelse(los > 7, 1, 0)) # Flag for prolonged ICU stays

## Save the combined dataset
write.csv(icu_wound_data, "icu_wound_data.csv", row.names = FALSE)
```
# Section 5: Preprocessing the Data
```{r}
# Step 1: Clean and Consolidate Data

# Identify and Remove Duplicates
## Check for duplicates
duplicates <- icu_wound_data %>%
  filter(duplicated(select(., subject_id, stay_id, hadm_id)))

## Remove duplicates
icu_wound_data <- icu_wound_data %>%
  distinct()
# Consolidate value.x and value.y
## Convert both columns to numeric
icu_wound_data <- icu_wound_data %>%
  mutate(
    value.x = as.numeric(value.x),
    value.y = as.numeric(value.y)
  )

icu_wound_data <- icu_wound_data %>%
  mutate(
    value.x = as.numeric(ifelse(value.x %in% c("N/A", ""), NA, value.x)),
    value.y = as.numeric(ifelse(value.y %in% c("N/A", ""), NA, value.y))
  )

icu_wound_data <- icu_wound_data %>%
  mutate(value = coalesce(value.x, value.y))
# Handle Missing Values
## Remove rows with missing values in critical columns
icu_wound_data <- icu_wound_data %>%
  filter(!is.na(value) & !is.na(los))

## Optionally, impute missing values for `value`
icu_wound_data$value <- ifelse(is.na(icu_wound_data$value), 
                               mean(icu_wound_data$value, na.rm = TRUE), 
                               icu_wound_data$value)
#  Drop Unnecessary Columns
icu_wound_data <- icu_wound_data %>%
  select(subject_id, stay_id, los, value, test_name, org_name, prolonged_stay)

# Summarize and Verify
## Check the structure of the cleaned data
str(icu_wound_data)

## View summary statistics
summary(icu_wound_data)

## Preview the first few rows
head(icu_wound_data)
```
# Section 6: Descriptive Statistics
```{r}
# Summarize ICU length of stay
summary(icu_wound_data$los)

# Summarize infection marker values
summary(icu_wound_data$value)

# Prolonged stay analysis
prolonged_summary <- icu_wound_data %>%
  group_by(prolonged_stay) %>%
  summarise(
    count = n(),
    mean_los = mean(los, na.rm = TRUE),
    mean_value = mean(value, na.rm = TRUE)
  )
print(prolonged_summary)

# Plot histogram of ICU length of stay
ggplot(icu_wound_data, aes(x = los)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Distribution of ICU Stay Duration",
       x = "Length of Stay (days)", y = "Frequency")

# Boxplot of infection marker values by prolonged stay
ggplot(icu_wound_data, aes(x = as.factor(prolonged_stay), y = value)) +
  geom_boxplot(fill = "orange") +
  labs(title = "Infection Marker Values by Prolonged Stay",
       x = "Prolonged Stay (0 = No, 1 = Yes)", y = "Infection Marker Value")
```
# Section 7: Regression Analysis
```{r}
# Linear regression: Predict value based on ICU stay duration and prolonged stay
reg_model <- lm(value ~ los + prolonged_stay, data = icu_wound_data)
summary(reg_model)

# Visualize regression results
ggplot(icu_wound_data, aes(x = los, y = value)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Regression: ICU Stay Duration vs Infection Marker Value",
       x = "Length of ICU Stay (days)", y = "Infection Marker Value")
```
# Section 8: Kaplan-Meier Survival Analysis
```{r}
# Simulate healing time based on `los` (replace this with actual healing data if available)
icu_wound_data <- icu_wound_data %>%
  mutate(healing_time = los + sample(1:5, n(), replace = TRUE), # Simulated healing time
         healed = ifelse(value < 5, 1, 0)) # Assume `value < 5` indicates healing

# Kaplan-Meier survival analysis
library(survival)

km_fit <- survfit(Surv(healing_time, healed) ~ prolonged_stay, data = icu_wound_data)

# Plot Kaplan-Meier survival curves
plot(km_fit, col = c("blue", "red"), lwd = 2,
     xlab = "Time to Wound Healing (days)",
     ylab = "Survival Probability",
     main = "Kaplan-Meier Curve: Time to Wound Healing")
legend("topright", legend = c("Short ICU Stay", "Prolonged ICU Stay"),
       col = c("blue", "red"), lwd = 2)
```
# Section 9: Random Forest Feature Importance
```{r}
# Ensure required library is installed
if (!require(randomForest)) install.packages("randomForest")
library(randomForest)

# Prepare data for random forest (remove NAs and select relevant variables)
rf_data <- icu_wound_data %>%
  select(value, los, prolonged_stay) %>%
  na.omit()

# Train random forest model
rf_model <- randomForest(
  value ~ .,  # Predict `value` using other variables
  data = rf_data,
  importance = TRUE,
  ntree = 500
)

# View feature importance
importance(rf_model)

# Plot feature importance
varImpPlot(rf_model)
```
# Section 10: Identifying Key Predictors of Impaired Healing
```{r}
# Check for the minimum and maximum values of the value column.
icu_wound_data %>% 
  summarise(
    min_value = min(value, na.rm = TRUE),
    max_value = max(value, na.rm = TRUE)
  )
# min_value: 0 max_value: 961
# Create a binary outcome for impaired healing using a threshold of 100
icu_wound_data <- icu_wound_data %>%
  mutate(impaired_healing = ifelse(value >= 100, 1, 0))

# Check the distribution of the new outcome variable
table(icu_wound_data$impaired_healing)

# Fit a logistic regression model to assess predictors of impaired healing
logit_model <- glm(impaired_healing ~ los + prolonged_stay, 
                   data = icu_wound_data, 
                   family = binomial)

# Display the summary of the model to review coefficient estimates and significance levels
summary(logit_model)

# Calculate odds ratios and 95% confidence intervals for the predictors
exp_coef <- exp(coef(logit_model))
conf_int <- exp(confint(logit_model))
odds_ratios <- data.frame(Estimate = exp_coef, conf_int)
print(odds_ratios)
```
# Section 11: Cox Proportional Hazards Model for Time to Healing
```{r}
# Ensure the survival library is loaded
library(survival)

# Fit the Cox proportional hazards model using healing time and event indicator (healed)
cox_model <- coxph(Surv(healing_time, healed) ~ los + prolonged_stay, 
                   data = icu_wound_data)

# Summarize the Cox model to review hazard ratios (HR) and p-values
summary(cox_model)
```